---
title: "Data compilation"
author: "Anne Marie Suffel"
date: "17 6 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(lubridate)
library(ggplot2)
library(rgdal)
library(broom)
require(maps)
library(sp)
library(magrittr)
```

loading the datasets

```{r message = FALSE}

getwd()
setwd("C:/Users/Anne M. Suffel/Documents/LSHTM/master thesis/Final data sets/raw_pop_data")
pop <- read.csv2("age_La1.csv", stringsAsFactors = FALSE)
fem <- read.csv2("female_LA1.csv", stringsAsFactors = FALSE)
mal <- read.csv2("male_LA1.csv", stringsAsFactors = FALSE)
eth <- read.csv2("ethnic_LA1.csv", stringsAsFactors = FALSE)
geo <- read.csv2("all_LA.csv", stringsAsFactors = FALSE)

head(pop)#code, Name, Geog, All_ages, age_0, ..., age_90
pop$code <- factor(pop$code)
pop$Name <- factor(pop$Name)
pop$Geog <- factor(pop$Geog)
summary(pop)


head(fem)#code, Name, Geography, All_ages, age_0, ..., age_90
fem$code <- factor(fem$code)
fem$Name <- factor(fem$Name)
fem$Geography <- factor(fem$Geography)

head(mal) #code, Name, Geography All_ages, age_0,..., age_90
mal$code <- factor(fem$code)
mal$Name <- factor(fem$Name)
mal$Geography <- factor(fem$Geography)

head(eth)#code, Area_Name, Total, White_British, Other_White, Mixed, Asian, Black, Other
eth$code <- factor(eth$code)
eth$Area_Name <- as.factor(eth$Area_Name)


head(geo) #code, area_names
geo$code <- factor(geo$code)
geo$area_names <- factor(geo$area_names)


```

Merging the datasets by LA Code for population per age, total of females, total of males and ethnicity


```{r message = FALSE}

df <- data.frame(code = geo$code, names = geo$area_names)
#summary(df)

#---merging with population per age
class(df$code)
class(pop$Code)
new_df <- df %>%
            dplyr::left_join(pop, by = "code") %>%
            dplyr::distinct ()
#merge.data.frame(df, pop, by = "code")
#summary(new_df)
head(new_df)


#---Merging with number of females in total
fem1 <- select(fem, code, All_ages)
#summary(fem1)
new_df <-new_df %>%
          dplyr::left_join(fem1, by = "code")
colnames(new_df)
head(new_df)
#r----enaming the new variables
new_df <- new_df %>%
            dplyr::rename(pop_tot = All_ages.x, fem_tot = All_ages.y)
#summary(new_df)
head(new_df)


#---mergin with number of males in total
mal1 <- select(mal, code, All_ages)
#summary(mal1)
new_df <- new_df %>%
            dplyr::left_join(mal1, by = "code") %>%
            dplyr::distinct()
#summary(new_df)
new_df <- rename(new_df, mal_tot = All_ages)
head(new_df)



#---merging with different ethnicities
#summary(eth)
#---correct number per thousand
eth1 <- eth %>%
          dplyr::mutate(White_British = (White_British * 1000), Other_White = (Other_White*1000), Mixed = (Mixed*1000),
                        Asian = (Asian*1000), Black = (Black*1000), Other = (Other*1000))
#summary(eth1)
eth1 <- select(eth1, code, White_British, Other_White, Mixed, Asian, Black, Other)
#summary(eth1)
#mergin data by LA
new_df <- new_df %>%
            dplyr::left_join (eth1, by = "code")
head(new_df)

#---deleting double variable
new_df <- select(new_df, -names)
#summary(new_df)

#---dropping unneccessary dataframes
remove(df, eth, eth1, fem, fem1, geo, mal, mal1, pop)
head(new_df)
      
````

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


Merging the new data frame with data about IMD

```{r message = FALSE}
getwd()
setwd("C:/Users/Anne M. Suffel/Documents/LSHTM/master thesis/Final data sets/raw_pop_data")
imd <- read.csv2("IMD_LA.csv",  stringsAsFactors = FALSE)

imd$IMD_score <- as.numeric(imd$IMD_score)
#summary(imd)
dim(imd) #317 rows and 22 columns
head(imd)
imd  <- rename(imd, code =  ï..code)


imd$code <- factor(imd$code)
imd$name <- factor(imd$name)
for(i in 3:22){
  imd[i] <- as.numeric(imd[,i])

}
#---check data transformation
head(imd)
new_imd <- imd %>%
  dplyr::select(code, name, IMD_rank)
new_imd <- new_imd %>%
            dplyr::mutate(imd_10 = IMD_rank)
new_imd <- new_imd %>%
              dplyr::mutate(imd_10 = ntile(IMD_rank, 10))


new_imd <- new_imd %>%
  dplyr::select(code, imd_10)

# #--calculation deciles for IMD overall
# new_imd <- new_imd %>%
#             dplyr::mutate(imd_10 = IMD_rank)
# new_imd <- new_imd %>%
#               dplyr::mutate(imd_10 = ntile(IMD_rank, 10))
# new_imd$imd_10 <- factor(new_imd$imd_10)
# #summary(new_imd)
# #---calculation for deciles for all the other imd parameters
# new_imd <- new_imd %>%
#               dplyr::mutate(income_10 = ntile(income_rank, 10), employement_10 = ntile(employement_rank, 10), education_10 = ntile(education_rank, 10),
#                              health_10 = ntile(health_rank, 10), crime_10 = ntile(crime_rank, 10), housing_10 = ntile(housing_rank, 10),
#                             environment_10 = ntile(environment_rank, 10), IDACI_10 = ntile(IDACI_rank, 10), IDAOPI_10 = ntile(IDAOPI_rank, 10))
# head(new_imd)                           
# #---transforming all the new deciles info factors
# 
# new_imd$income_10 <- factor(new_imd$income_10)
# new_imd$employement_10 <- factor(new_imd$employement_10)
# new_imd$education_10 <- factor(new_imd$education_10)
# new_imd$health_10 <- factor(new_imd$health_10)
# new_imd$crime_10 <- factor(new_imd$crime_10)
# new_imd$housing_10 <- factor(new_imd$housing_10)
# new_imd$environment_10 <- factor(new_imd$environment_10)
# new_imd$IDACI_10 <- factor(new_imd$IDACI_10)
# new_imd$IDAOPI_10 <- factor(new_imd$IDAOPI_10)
# 
# #--selecting the important stuff
# imd_merge <- select(new_imd, code, name, imd_10, income_10, employement_10, education_10, health_10, crime_10, housing_10, environment_10, IDACI_10, IDAOPI_10)
# #summary(imd_merge)

##Merging the population data frame with imd_deciles data frame
data_pop <- new_df %>%
              dplyr::left_join(new_imd, by = "code") %>%
              dplyr::distinct()
colnames(data_pop)


#cleaning working environment
remove( new_df, new_imd)

````



exporting newly created data frame without comorbidities
````{r message = FALSE}

save(data_pop, file = "C:/Users/Anne M. Suffel/Documents/LSHTM/master thesis/Final data sets/pop.Rdata")

````



Attach the data about comorbidities as well

```{r message = FALSE}

#pop1 <- load(file = "C:/Users/Anne M. Suffel/Documents/LSHTM/master thesis/Final data sets/pop.Rdata")


##now load comorbidties for ONS population
setwd("C:/Users/Anne M. Suffel/Documents/LSHTM/master thesis/Final data sets/raw_comorb")
obesity <- read.csv2("obs_LA.csv", stringsAsFactors = FALSE)
diabetes <- read.csv2("diab_LA.csv", stringsAsFactors = FALSE)
hyper <- read.csv2("hyper_LA.csv", stringsAsFactors = FALSE)


##data formatting for merging; forming accicental factor variables into numerics
head(obesity)
obesity <- dplyr::rename(obesity, code = ï..code)
obesity$obs_tot <- as.numeric(obesity$obs_tot)  ##absolute number in the LA 
obesity$code <- factor(obesity$code)
obesity$name <- factor(obesity$name)
head(obesity)


head(hyper)
hyper$hyp_total <- as.numeric(hyper$hyp_total) #total numer of Hypertension
hyper$hyp_prev <- as.numeric(hyper$hyp_prev)#percentage of hypertension
hyper$code <- factor(hyper$code)
hyper$name <- factor(hyper$name)
head(hyper)

head(diabetes)
diabetes$total_diab <- as.numeric(diabetes$total_diab)
diabetes$diab_prev_perc <- as.numeric(diabetes$diab_prev_perc)
diabetes$diab_prev_perc <- as.numeric(diabetes$diab_prev_perc)
diabetes$diab_prev <- as.numeric(diabetes$diab_prev)
diabetes$code <- factor(diabetes$code)
diabetes$name <- factor(diabetes$name)
head(diabetes)


````
 Merginng the population data set with the different comorbidity data_sets

````{r message = FALSE}
data_comorb <- data_pop %>%
                dplyr::left_join(obesity, by = "code") %>%
                dplyr::distinct()

data_comorb <- data_comorb %>%
                dplyr::left_join(diabetes, by = "code") %>%
                dplyr::distinct()

data_comorb <- data_comorb %>%
                dplyr::left_join(hyper, by = "code") %>%
                dplyr::distinct()

head(data_comorb)



##now calculate only absolute number of cases and then rate per 100,000
data_comorb <- data_comorb %>%
                dplyr::select(-c(name.x, , name, name.y, diab_prev_perc, hyp_prev, diab_prev))


colnames(data_comorb)
data_comorb <- data_comorb %>%
                  dplyr::rename(diab_tot = "total_diab", hyper_tot = "hyp_total")


data_comorb <- data_comorb %>%
                  dplyr::mutate(obs_ratio = ((obs_tot * 100000)/pop_tot), diab_ratio = ((diab_tot * 100000)/pop_tot), 
                                hyper_ratio = ((hyper_tot * 100000)/pop_tot))
head(data_comorb)


##saving data file
save(data_comorb, file = "C:/Users/Anne M. Suffel/Documents/LSHTM/master thesis/Final data sets/pop_comorb.Rdata")

````



